---
title: "TA Session 1"
author: "Chi-Yuan Fang"
date: "March 2, 2021"
output: 
  html_document:
    toc: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#library(kableExtra)
```

# Introduction

## TA Information

TA: Chi-Yuan Fang

TA sessions: Tuesday 1:20 – 3:10 PM (SS 501) 

Email: r09323017@ntu.edu.tw

Office hours: Tuesday 10:00 - 11:30 AM or by appointments (SS 643)

Class group on Facebook: Statistics (Fall 2020) and Econometrics (Spring 2021) <https://www.facebook.com/groups/452292659024369/>

Because screens are not clear in SS 501, I would provide live screen in the group.


## TA Sessions Schedule

| Week | TA Sessions    | Quiz | Content | Remind |
|------|-------------|------|-------------------------|-----------------|
|  1   | 02/23: No class| | | | |
|  2   | 03/02: Class 1 | | Function, Confidence Interval, T test | 03/10 Turn in HW1 |
|  3   | 03/09: Class 2 | |  | 03/10 Turn in HW1, 03/16 Quiz 1 |
|  4   | 03/16: Class 3 | Quiz 1 |  | 03/24 Turn in HW2 |
|  5   | 03/23: Class 4 | |  | 03/24 Turn in HW2, 03/30 Quiz 2 |
|  6   | 03/30: Class 5 | Quiz 2 |  | 04/14 Turn in HW3 |
|  7   | 04/06: No class | | | 04/14 Turn in HW3 |
|  8   | 04/13: Class 6 | |  | 04/14 Turn in HW3, 04/20 Quiz 3 |
|  9   | 04/20: Class 7 | Quiz 3 | | **04/28 Midterm** |
| 10   | 04/27: Class 8 | | Review and Q&A | **04/28 Midterm**, 05/05 Turn in HW4 |
| 11   | 05/04: Class 9| |  | 05/05 Turn in HW4, 05/11 Quiz 4 |
| 12   | 05/11: Class 10| Quiz 4 |  | 05/19 Turn in HW5 |
| 13   | 05/18: Class 11| |  | 05/19 Turn in HW5, 05/25 Quiz 5 |
| 14   | 05/25: Class 12| Quiz 5 |  | 06/02 Turn in HW6 |
| 15   | 06/01: Class 13| |  | 06/02 Turn in HW6, 06/08 Quiz 6 |
| 16   | 06/08: Class 14| Quiz 6 | Review and Q&A | **06/16 Final Exam** | 
| 17   | 06/15: No class| | | **06/16 Final Exam** | 
| 18   | 06/22: No class| | |  |

## Reference

What is a good book on learning R with examples? <https://www.quora.com/What-is-a-good-book-on-learning-R-with-examples>

# Empircial Exercise 3.1

> On the text website, <http://www.pearsonglobaleditions.com>, you will find the data file **CPS96_15**, which contains an extended version of the data set used in Table 3.1 of the text for the years 1996 and 2015. It contains data on full-time
workers, ages 25–34, with a high school diploma or a B.A./B.S. as their highest
degree. A detailed description is given in **CPS96_15_Description**, available on
the website. Use these data to complete the following.

> a.
>      i. Compute the sample mean for average hourly earnings ($AHE$) in 1996 and 2015.
>      ii. Compute the sample standard deviation for $AHE$ in 1996 and 2015.
>      iii. Construct a 95\% confidence interval for the population means of $AHE$
in 1996 and 2015.
>      iv. Construct a 95\% confidence interval for the change in the population
means of $AHE$ between 1996 and 2015.

**Solution**

```{r}
# import data 
library(readxl)
CPS96_15 <- read_xlsx("CPS96_15/CPS96_15.xlsx")

# 1996
CPS96 <- CPS96_15[CPS96_15$year == 1996,]

# 2015
CPS15 <- CPS96_15[CPS96_15$year == 2015,]

# a. # i.-ii
E31a <- function(x){
  
  # i. mean
  mu <- mean(x)
  
  # ii. standard error
  se <- sd(x)/length(x)
  
  # test
  test <- t.test(x, 
                 alternative = c("two.sided"),
                 mu = 0, # H0
                 conf.level = 0.95) # alpha = 0.05
  
  # iii. CI
  lower <- test$conf.int[1]
  upper <- test$conf.int[2]
  CI <- paste(lower, "-"  ,upper)
  
  Table <- data.frame(mu, se, CI)
  colnames(Table) <- c("Mean", "Standard Error", "95% Confidence Interval")
  
  Table

}

tapply(CPS96_15$ahe, CPS96_15$year, E31a)

# a. # iv.

# Difference
t.test(CPS15$ahe, CPS96$ahe, 
       alternative = c("two.sided"),
       mu = 0.5, # H0
       var.equal = FALSE,
       conf.level = 0.95) # alpha = 0.05

```

> b. In 2015, the value of the Consumer Price Index (CPI) was 237.0. In 1996, the value of the CPI was 156.9. Repeat (a), but use AHE measured in real 2015 dollars (\$2015); that is, adjust the 1996 data for the price inflation that occurred between 1996 and 2015.

**Solution**



> c. If you were interested in the change in workers’ purchasing power from 1996 to 2015, would you use the results from (a) or (b)? Explain.

**Solution**

> d. Using the data for 2015:
>      i. Construct a 95\% confidence interval for the mean of $AHE$ for high school graduates. 
>      ii. Construct a 95\% confidence interval for the mean of $AHE$ for
workers with a college degree.
>      iii. Construct a 95\% confidence interval for the difference between the
two means.

**Solution**

> e. Repeat (d) using the 1996 data expressed in \$2015.

**Solution**

> f. Using appropriate estimates, confidence intervals, and test statistics, answer the following questions:
>      i. Did real (inflation-adjusted) wages of high school graduates increase from 1996 to 2015?
>      ii. Did real wages of college graduates increase?
>      iii. Did the gap between earnings of college and high school graduates increase? Explain.

**Solution**

> g. Table 3.1 presents information on the gender gap for college graduates. Prepare a similar table for high school graduates, using the 1996 and 2015 data. Are there any notable differences between the results for high school and college graduates?

**Solution**



