---
title: "Empirical Exercise - E5.3"
author: "Chi-Yuan Fang"
date: "`r Sys.Date()`"
output:
  html_document:
    number_sections: yes
    toc: yes
  pdf_document:
    toc: yes

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#library(kableExtra)
```

```{r packages, include=FALSE}
library(openintro)
library(tidyverse)
data(COL)
```

> On the text website, <http://www.pearsonglobaleditions.com>, you will find the data file **Birthweight_Smoking**, which contains data for a random sample of babies born in Pennsylvania in 1989. The data include the baby’s birth weight together with various characteristics of the mother, including whether she smoked during the pregnancy. A detailed description is given in **Birthweight_Smoking_Description**, also available on the website. In this exercise, you will investigate the relationship between birth weight and smoking during pregnancy.

> a. In the sample:
>     i. What is the average value of $Birthweight$ for all mothers?
>     ii. For mothers who smoke?
>     iii. For mothers who do not smoke?

**Solution**

```{r}
# import data 
library(readxl)
Birthweight_Smoking <- read_xlsx("Birthweight_Smoking/Birthweight_Smoking.xlsx")

# i 
mean(Birthweight_Smoking$birthweight)

# ii. # iii.
tapply(Birthweight_Smoking$birthweight, Birthweight_Smoking$smoker, mean)


```

> b. 
>     i. Use the data in the sample to estimate the difference in average birth weight for smoking and nonsmoking mothers.
>     ii. What is the standard error for the estimated difference in (i)?
>     iii. Construct a 95\% confidence interval for the difference in the average birth weight for smoking and nonsmoking mothers.

**Solution**

> c. Run a regression of $Birthweight$ on the binary variable $Smoker$.
>     i. Explain how the estimated slope and intercept are related to your answers in parts (a) and (b).
>     ii. Explain how the $SE\left( \beta_1 \right)$ is related to your answer in b(ii).
>     iii. Construct a 95\% confidence interval for the effect of smoking on birth weight.

**Solution**

i. The intercept is the average birthweight for non-smokers ($Smoker = 0$). The slope is the difference between average birthweights for smokers ($Smoker = 1$) and non-smokers ($Smoker = 0$).

ii. They are the same.

iii. This is same as in b(iii).


> d. Do you think smoking is uncorrelated with other factors that cause
low birth weight? That is, do you think that the regression error term—say, $u_i$—has a conditional mean of 0 given $Smoking$ $(X_i)$? (You will investigate this further in $Birthweight$ and $Smoking$ exercises in later chapters.)

**Solution**



