---
title: "Empirical Exercise - E5.1"
author: "Chi-Yuan Fang"
date: "`r Sys.Date()`"
output:
  html_document:
    number_sections: yes
    toc: yes
  pdf_document:
    toc: yes

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#library(kableExtra)
```

```{r packages, include=FALSE}
library(openintro)
library(tidyverse)
data(COL)
```

> On the text website, <http://www.pearsonglobaleditions.com>, you will find the data file **CPS96_15**, which contains an extended version of the data set used in Table 3.1 of the text for the years 1996 and 2015. It contains data on full-time workers, ages 25–34, with a high school diploma or a B.A./B.S. as their highest degree. A detailed description is given in **CPS96_15_Description**, available on the website. Use these data to complete the following.

> a.
>      i. Compute the sample mean for average hourly earnings ($AHE$) in 1996 and 2015.
>      ii. Compute the sample standard deviation for $AHE$ in 1996 and 2015.
>      iii. Construct a 95\% confidence interval for the population means of $AHE$ in 1996 and 2015.
>      iv. Construct a 95\% confidence interval for the change in the population means of $AHE$ between 1996 and 2015.

**Solution**

```{r}
# import data 
library(readxl)
CPS96_15 <- read_xlsx("CPS96_15/CPS96_15.xlsx")

# 1996
CPS96 <- CPS96_15[CPS96_15$year == 1996,]

# 2015
CPS15 <- CPS96_15[CPS96_15$year == 2015,]

# a. # i.-ii
E31a <- function(x){
  # i. sample mean
  mu <- mean(x)
  
  # ii. sample standard deviation (standard error)
  se <- sd(x)/sqrt(length(x))
  
  # test
  test <- t.test(x, 
                 alternative = c("two.sided"),
                 mu = 0, # H0
                 conf.level = 0.95) # alpha = 0.05
  
  # iii. 95% confidence interval
  lower <- round(test$conf.int[1], digit = 4)
  upper <- round(test$conf.int[2], digit = 4)
  CI <- paste(lower, "-"  ,upper)
  
  Table <- data.frame(mu, se, CI)
  colnames(Table) <- c("Mean", "Standard Error", "95% Confidence Interval")
  
  Table

}

tapply(CPS96_15$ahe, CPS96_15$year, E31a)

# a. # iv. 95% confidence interval for difference

t.test(CPS15$ahe, CPS96$ahe, 
       alternative = c("two.sided"),
       mu = 0, # H0
       var.equal = FALSE,
       conf.level = 0.95) # alpha = 0.05


```

> b. In 2015, the value of the Consumer Price Index (CPI) was 237.0. In 1996, the value of the CPI was 156.9. Repeat (a), but use AHE measured in real 2015 dollars (\$2015); that is, adjust the 1996 data for the price inflation that occurred between 1996 and 2015.

**Solution**

```{r}
# 1996 data
CPS96 <- CPS96_15[CPS96_15$year == 1996,]

# CPI
CPI_96 <- 156.9
CPI_15 <- 237

# adjusted 1996 AHE in $2015 
ahe_adjust <- CPS96$ahe * (CPI_15/CPI_96)

CPS96 <- cbind(CPS96, ahe_adjust)

# i. sample mean
# ii. sample standard deviation
# iii. 95% confidence interval
# for adjusted 1996 AHE in $2015 
E31a(CPS96$ahe_adjust)

# iv. 95% confidence interval for difference
t.test(CPS15$ahe, CPS96$ahe_adjust, 
       alternative = c("two.sided"),
       mu = 0, # H0
       var.equal = FALSE,
       conf.level = 0.95) # alpha = 0.05

```

> c. If you were interested in the change in workers’ purchasing power from 1996 to 2015, would you use the results from (a) or (b)? Explain.

**Solution**

The results from part (b) adjust for changes in purchasing power. These results should be
used.

> d. Using the data for 2015:
>      i. Construct a 95\% confidence interval for the mean of $AHE$ for high school graduates. 
>      ii. Construct a 95\% confidence interval for the mean of $AHE$ for
workers with a college degree.
>      iii. Construct a 95\% confidence interval for the difference between the
two means.

**Solution**

```{r}
# high school in 2015
CPS15hs <- CPS15[CPS15$bachelor == 0, ]

# college in 2015
CPS15col <- CPS15[CPS15$bachelor == 1, ]

# i. 95% CI for AHE in high school
# ii. 95% CI for AHE in college
tapply(CPS15$ahe, CPS15$bachelor, E31a)

# iii. 95% CI for difference
t.test(CPS15hs$ahe, CPS15col$ahe, 
       alternative = c("two.sided"),
       mu = 0, # H0
       var.equal = FALSE,
       conf.level = 0.95) # alpha = 0.05

```


> e. Repeat (d) using the 1996 data expressed in \$2015.

**Solution**

```{r}
# high school in 2015
CPS96hs <- CPS96[CPS96$bachelor == 0, ]

# college in 2015
CPS96col <- CPS96[CPS96$bachelor == 1, ]

# i. # ii.
tapply(CPS96$ahe_adjust, CPS96$bachelor, E31a)

# iii. 95% CI for difference
t.test(CPS96col$ahe_adjust, CPS96hs$ahe_adjust, 
       alternative = c("two.sided"),
       mu = 0, # H0
       var.equal = FALSE,
       conf.level = 0.95) # alpha = 0.05

```


> f. Using appropriate estimates, confidence intervals, and test statistics, answer the following questions:
>      i. Did real (inflation-adjusted) wages of high school graduates increase from 1996 to 2015?
>      ii. Did real wages of college graduates increase?
>      iii. Did the gap between earnings of college and high school graduates increase? Explain.

**Solution**

i. 
*  **Prepare**

$H_0: wage_{2015}^{hs} - wage_{1996}^{hs} = 0$ v.s. $H_1: wage_{2015}^{hs} - wage_{1996}^{hs} \neq 0$

Let the significance level be 0.05.


*  **Calculate**

```{r}
t.test(CPS15hs$ahe, CPS96hs$ahe_adjust, 
       alternative = c("two.sided"),
       mu = 0, # H0
       var.equal = FALSE,
       conf.level = 0.95) # alpha = 0.05

```

*  **Conclude**

Because $p-value > 0.05$, we reject $H_0$. There is no evidence that real (inflation-adjusted) wages of high school graduates increase from 1996 to 2015.

ii.

*  **Prepare**

$H_0: wage_{2015}^{col} - wage_{1996}^{col} = 0$ v.s. $H_1: wage_{2015}^{col} - wage_{1996}^{col} \neq 0$

Let the significance level be 0.05.


*  **Calculate**

```{r}
t.test(CPS15col$ahe, CPS96col$ahe_adjust, 
       alternative = c("two.sided"),
       mu = 0, # H0
       var.equal = FALSE,
       conf.level = 0.95) # alpha = 0.05

```

*  **Conclude**

Because $p-value < 0.05$, we reject $H_0$. There is statistically significance evidence that real (inflation-adjusted) wages of college graduates increase from 1996 to 2015.


iii. (Skip.)

*  **Prepare**

$H_0: wage_{2015}^{diff} - wage_{1996}^{diff} = 0$ v.s. $H_1: wage_{2015}^{diff} - wage_{1996}^{diff} \neq 0$

Let the significance level be 0.05.


*  **Calculate**

```{r}
t.test(CPS96col$ahe_adjust, CPS96hs$ahe_adjust, 
       alternative = c("two.sided"),
       mu = 0, # H0
       var.equal = FALSE,
       conf.level = 0.95) # alpha = 0.05

t.test(CPS15col$ahe, CPS15hs$ahe, 
       alternative = c("two.sided"),
       mu = 0, # H0
       var.equal = FALSE,
       conf.level = 0.95) # alpha = 0.05

```

*  **Conclude**

Skip.


> g. Table 3.1 presents information on the gender gap for college graduates. Prepare a similar table for high school graduates, using the 1996 and 2015 data. Are there any notable differences between the results for high school and college graduates?

**Solution**

```{r}
E31g <- function(x){
  # sample mean
  mu <- mean(x)
  
  # standard deviation
  SD <- sd(x)
  
  # numbers of sample
  n <- length(x)
  
  Table <- data.frame(mu, SD, n)
  
  Table
}

# data: male in 1996
CPS96hsm <- CPS96hs[CPS96hs$female == 0, ]

# data: female in 1996
CPS96hsf <- CPS96hs[CPS96hs$female == 1, ]

# statistics: male and female in 1996
tapply(CPS96hs$ahe_adjust, CPS96hs$female, E31g)

# difference in 1996
t.test(CPS96hsm$ahe_adjust, CPS96hsf$ahe_adjust, 
       alternative = c("two.sided"),
       mu = 0, # H0
       var.equal = FALSE,
       conf.level = 0.95) # alpha = 0.05

# statistics: male and female in 2015
tapply(CPS15hs$ahe, CPS15hs$female, E31g)

# data: male in 2015
CPS15hsm <- CPS15hs[CPS15hs$female == 0, ]

# data: female in 2015
CPS15hsf <- CPS15hs[CPS15hs$female == 1, ]

# difference in 2015
t.test(CPS15hsm$ahe, CPS15hsf$ahe, 
       alternative = c("two.sided"),
       mu = 0, # H0
       var.equal = FALSE,
       conf.level = 0.95) # alpha = 0.05

```

```{r}
library(alfred)

indpro <- get_fred_series("INDPRO", "indpro")
```

