---
title: "Empirical Exercise - E5.2"
author: "Chi-Yuan Fang"
date: "`r Sys.Date()`"
output:
  html_document:
    number_sections: yes
    toc: yes
  pdf_document:
    toc: yes

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#library(kableExtra)
```

```{r packages, include=FALSE}
library(openintro)
library(tidyverse)
data(COL)
```

> Using the data set *Growth* described in Empirical Exercise 4.1, but excluding the data for Malta, run a regression of $Growth$ on $TradeShare$.

> a. Is the estimated regression slope statistically significant? That is, can you reject the null hypothesis $H_0: \beta_1 = 0$ vs. a two-sided alternative hypothesis at the 10\%, 5\%, or 1\% significance level?

**Solution**

*  **Prepare**

$H_0: \beta_{height} = 0$ v.s. $H_1: \beta_{height} \neq 0$

Let the significance level be 0.10, 0.05, or 0.01.

*  **Calculate**

```{r}
# import data 
library(readxl)
Growth <- read_xlsx("Growth/Growth.xlsx")

Growth_n <- Growth[Growth$country_name != "Malta",]

E52a_model <- lm(growth ~ tradeshare, data = Growth_n)

summary(E52a_model)

```

*  **Conclude**

Because $p-value = 0.0937$, we reject $H_0$ at $\alpha = 0.10$, but do not reject $H_0$ at $\alpha = 0.05 \mbox{ or } 0.01$. 

> b. What is the $p$-value associated with the coefficient’s $t$-statistic?

**Solution**

See part (a).

> c. Construct a 90\% confidence interval for $\beta_1$.

**Solution**

```{r}
E52c <- function(x, y){
  # numbers of sample
  n <- length(y)
  
  # sample mean
  xbar <- mean(x)
  ybar <- mean(y)
  
  # OLS coefficient
  b1hat <- cov(x,y)/var(x)
  b0hat <- ybar - b1hat*xbar
  
  yhat <- b0hat + b1hat*x
  
  # explained sum of squares (ESS)
  ESS <- sum((yhat - ybar)^2)
  # total sum of squares (TSS)
  TSS <- sum((y - ybar)^2)
  # sum of squared residuals (SSR)
  SSR <- sum((y - yhat)^2)
  
  # coefficient of determination
  #Rsqure <- ESS/TSS
  
  # standard error of the regression
  SER <- sqrt(SSR/(n-2))
  
  # standard error of coefficient
  se_b1hat <- sqrt(SER^2/sum((x - xbar)^2))
  
  # 95% CI for b1hat
  lower_b1hat <- round(b1hat - qnorm(0.95, mean = 0, sd = 1)*se_b1hat, digit = 4)
  upper_b1hat <- round(b1hat + qnorm(0.95, mean = 0, sd = 1)*se_b1hat, digit = 4)
  CI_b1hat <- paste(lower_b1hat, "-"  ,upper_b1hat)
  
  Table <- matrix(c(b1hat, se_b1hat, CI_b1hat), nrow = 1)
  colnames(Table) <- c("Estimate", "Standard Error", "90% Confidence Interval")
  
  Table
} 

E52c(Growth_n$tradeshare, Growth_n$growth)
```




