---
title: "TA Session 1"
author: "Chi-Yuan Fang"
date: "March 2, 2021"
output: 
  html_document:
    toc: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#library(kableExtra)
```

# Introduction

## TA Information

TA: Chi-Yuan Fang

TA sessions: Tuesday 1:20 – 3:10 PM (SS 501) 

Email: r09323017@ntu.edu.tw

Office hours: Friday 14:00 - 15:30 AM or by appointments (SS 643)

Class group on Facebook: Statistics (Fall 2020) and Econometrics (Spring 2021) <https://www.facebook.com/groups/452292659024369/>

Because screens are not clear in SS 501, I will provide live screen in the group.


## TA Sessions Schedule

| Week | TA Sessions    | Quiz | Content | Remind |
|------|-------------|------|-------------------------|-----------------|
|  1   | 02/23: No class| | | | |
|  2   | 03/02: Class 1 | | Function, Confidence Interval, T test | 03/10 Turn in HW1 |
|  3   | 03/09: Class 2 | |  | 03/10 Turn in HW1, 03/16 Quiz 1 |
|  4   | 03/16: Class 3 | Quiz 1 |  | 03/24 Turn in HW2 |
|  5   | 03/23: Class 4 | |  | 03/24 Turn in HW2, 03/30 Quiz 2 |
|  6   | 03/30: Class 5 | Quiz 2 |  | 04/14 Turn in HW3 |
|  7   | 04/06: No class | | | 04/14 Turn in HW3 |
|  8   | 04/13: Class 6 | |  | 04/14 Turn in HW3, 04/20 Quiz 3 |
|  9   | 04/20: Class 7 | Quiz 3 | | **04/28 Midterm** |
| 10   | 04/27: Class 8 | | Review and Q&A | **04/28 Midterm**, 05/05 Turn in HW4 |
| 11   | 05/04: Class 9| |  | 05/05 Turn in HW4, 05/11 Quiz 4 |
| 12   | 05/11: Class 10| Quiz 4 |  | 05/19 Turn in HW5 |
| 13   | 05/18: Class 11| |  | 05/19 Turn in HW5, 05/25 Quiz 5 |
| 14   | 05/25: Class 12| Quiz 5 |  | 06/02 Turn in HW6 |
| 15   | 06/01: Class 13| |  | 06/02 Turn in HW6, 06/08 Quiz 6 |
| 16   | 06/08: Class 14| Quiz 6 | Review and Q&A | **06/16 Final Exam** | 
| 17   | 06/15: No class| | | **06/16 Final Exam** | 
| 18   | 06/22: No class| | |  |

## Reference

Introduction to Econometrics with R

<https://www.econometrics-with-r.org>

R for Data Science 

<https://r4ds.had.co.nz>

R Markdown

<https://rmarkdown.rstudio.com>

Introduction to R Markdown

<https://rpubs.com/brandonkopp/RMarkdown>

What is a good book on learning R with examples?

<https://www.quora.com/What-is-a-good-book-on-learning-R-with-examples>

# Large-Sample Approximations to Sampling Distributions

## The Law of Large Numbers and Consistency

### Figure 2.8 [p.87]

Figure 2.8: Sampling Distribution of the Sample Average of $n$ Bernoulli Random Variables [p.87]

```{r}
# p.87 Figure 2.8 

# set sample size and number of samples
Figure28 <- function(n, mu, reps, index){

  # n x reps sample matrix
  samples <- replicate(reps, rbinom(n, size = 1, prob = mu)) 
  
  # compute sample mean
  sample.avgs <- colMeans(samples)
  
  # histogram
  hist(sample.avgs, 
       probability = TRUE, # freq = FALSE
       col = "steelblue",
       xlim = c(0, 1),
       xlab = "Value of sample average",
       main = paste("(", index, ") n =", n)) # 
}

# subdivide the plot panel into a 2-by-2 array
# https://www.r-bloggers.com/2012/06/two-tips-adding-title-for-graph-with-multiple-plots-add-significance-asterix-onto-a-boxplot/
par(mfrow=c(2, 2),oma = c(0, 0, 2, 2))

# set the number of repetitions and the sample sizes
reps <- 10000
sample.sizes <- c(2, 5, 25, 100)

# set seed for reproducibility
set.seed(12345)

for (i in sample.sizes){
  Figure28(i, 0.78, reps, letters[which(sample.sizes == i)])
}

mtext("Figure 2.8", outer = TRUE, cex = 1.3, font = 2)

```



## The Central Limit Theorem

### Figure 2.9 [p.88]

Figure 2.9: Distribution of the Standardized Sample Average of $n$ Bernoulli Random Variables with $p = 0.78$ [p.88]

```{r}
# p.88 Figure 2.9

Figure29 <- function(n, mu, reps, index){
  # initialize the vector of sample means
  samplemean <- rep(0, reps) # 0 for reps times
  
  # initialize the vector of standardized sample means
  stdsamplemean <- rep(0, reps) # 0 for reps times
  
  # inner loop (loop over repetitions)   
  for (i in 1:reps){
    # generate sample
    x <- rbinom(n, 1, mu) # mu = 0.78
    
    # sample mean
    samplemean[i] <- mean(x)
    
    # standardized
    stdsamplemean[i] <- (mean(x) - mu)/sqrt(mu*(1-mu)/n) 
  }
  
  # plot histogram
  hist(stdsamplemean,
       probability = TRUE, # freq = FALSE
       col = "steelblue",
       breaks = 40,
       xlim = c(-3, 3), ylim = c(0, 0.8),
       xlab = "Standardized value of sample average", 
       main = paste("(", index, ") n = ", n))
  
  # overlay the N(0,1) density for every iteration    
  curve(dnorm(x), lwd = 2, col = "black", add = TRUE)  

}

# subdivide the plot panel into a 2-by-2 array
# https://www.r-bloggers.com/2012/06/two-tips-adding-title-for-graph-with-multiple-plots-add-significance-asterix-onto-a-boxplot/
par(mfrow=c(2, 2),oma = c(0, 0, 2, 2))

# set the number of repetitions and the sample sizes
reps <- 10000
sample.sizes <- c(2, 5, 25, 100)

# set seed for reproducibility
set.seed(12345)
  
for (i in sample.sizes){
  Figure29(i, 0.78, reps, letters[which(sample.sizes == i)])
}

mtext("Figure 2.9", outer = TRUE, cex = 1.3, font = 2)
 
```

### Figure 2.10 [p.90]

Figure 2.10: Distribution of the Standardized Sample Average of $n$ Draws from a Skewed Population Distribution [p.90]

```{r}
# p.90 Figure 2.10

# https://stackoverflow.com/questions/4807398/how-to-generate-distributions-given-mean-sd-skew-and-kurtosis-in-r

# https://stackoverflow.com/questions/1497539/fitting-a-density-curve-to-a-histogram-in-r

library("PearsonDS")

Figure210 <- function(n, mu, var, skew, kurt, reps, index){
  # set moments
  moments <- c(mean = mu, variance = var, skewness = skew, kurtosis = kurt)
 
  # initialize the vector of sample means
  samplemean <- rep(0, reps) # 0 for reps times
  
  # initialize the vector of standardized sample means
  stdsamplemean <- rep(0, reps) # 0 for reps times

  # inner loop (loop over repetitions)   
  for (i in 1:reps){
    # generate sample
    x <- rpearson(reps, moments = moments)
    
    # sample mean
    samplemean[i] <- mean(x)
    
    # standardized
    stdsamplemean[i] <- (mean(x) - mu)/sqrt(var/n) 
  }    
  
  # plot histogram
  hist(stdsamplemean,
       probability = TRUE, # freq = FALSE
       col = "steelblue",
       breaks = 40,
       xlim = c(-3, 3), #ylim = c(0, 0.8),
       xlab = "Standardized value of sample average", 
       main = paste("(", index, ") n = ", n))
  
  # overlay the N(0,1) density for every iteration    
  curve(dnorm(x), lwd = 2, col = "black", add = TRUE)  
  
}

# subdivide the plot panel into a 3-by-2 array
# https://www.r-bloggers.com/2012/06/two-tips-adding-title-for-graph-with-multiple-plots-add-significance-asterix-onto-a-boxplot/
par(mfrow=c(3, 2), oma = c(0, 0, 3, 3))

# set the number of repetitions and the sample sizes
reps <- 1000
sample.sizes <- c(1, 5, 25, 100, 500, 1000)

# set seed for reproducibility
set.seed(12345)

for (i in sample.sizes){
  Figure210(i, 0, 1, 0.1, 3, reps, letters[which(sample.sizes == i)])
}

mtext("Figure 2.10", outer = TRUE, cex = 1.3, font = 2)

```


# Empircial Exercise 3.1

> On the text website, <http://www.pearsonglobaleditions.com>, you will find the data file **CPS96_15**, which contains an extended version of the data set used in Table 3.1 of the text for the years 1996 and 2015. It contains data on full-time workers, ages 25–34, with a high school diploma or a B.A./B.S. as their highest degree. A detailed description is given in **CPS96_15_Description**, available on the website. Use these data to complete the following.

> a.
>      i. Compute the sample mean for average hourly earnings ($AHE$) in 1996 and 2015.
>      ii. Compute the sample standard deviation for $AHE$ in 1996 and 2015.
>      iii. Construct a 95\% confidence interval for the population means of $AHE$ in 1996 and 2015.
>      iv. Construct a 95\% confidence interval for the change in the population means of $AHE$ between 1996 and 2015.

**Solution**

```{r}
# import data 
library(readxl)
CPS96_15 <- read_xlsx("CPS96_15/CPS96_15.xlsx")

# data: 1996
CPS96 <- CPS96_15[CPS96_15$year == 1996,]

# data: 2015
CPS15 <- CPS96_15[CPS96_15$year == 2015,]

# i. sample mean # ii. standard error # iii. 95% CI
E31a <- function(x){
  # i. sample mean
  mu <- mean(x)
  
  # ii. sample standard deviation (standard error)
  se <- sd(x)/sqrt(length(x))
  
  # test
  test <- t.test(x, 
                 alternative = c("two.sided"),
                 mu = 0, # H0
                 conf.level = 0.95) # alpha = 0.05
  
  # iii. 95% confidence interval
  lower <- round(test$conf.int[1], digit = 4)
  upper <- round(test$conf.int[2], digit = 4)
  CI <- paste(lower, "-"  ,upper)
  
  Table <- data.frame(mu, se, CI)
  colnames(Table) <- c("Mean", "Standard Error", "95% Confidence Interval")
  
  Table

}

tapply(CPS96_15$ahe, CPS96_15$year, E31a)

# iv. 95% confidence interval for difference

t.test(CPS15$ahe, CPS96$ahe, 
       alternative = c("two.sided"),
       mu = 0, # H0
       var.equal = FALSE,
       conf.level = 0.95) # alpha = 0.05

```

> b. In 2015, the value of the Consumer Price Index (CPI) was 237.0. In 1996, the value of the CPI was 156.9. Repeat (a), but use AHE measured in real 2015 dollars (\$2015); that is, adjust the 1996 data for the price inflation that occurred between 1996 and 2015.

**Solution**

```{r}
# data: 1996
CPS96 <- CPS96_15[CPS96_15$year == 1996,]

# CPI
CPI_96 <- 156.9
CPI_15 <- 237

# adjusted 1996 AHE in $2015 
ahe_adjust <- CPS96$ahe * (CPI_15/CPI_96)

# data: 1996 including adjusted 1996 AHE in $2015 
CPS96 <- cbind(CPS96, ahe_adjust)

# i. sample mean
# ii. sample standard deviation
# iii. 95% confidence interval
# for adjusted 1996 AHE in $2015 
E31a(CPS96$ahe_adjust)

# iv. 95% confidence interval for difference
t.test(CPS15$ahe, CPS96$ahe_adjust, 
       alternative = c("two.sided"),
       mu = 0, # H0
       var.equal = FALSE,
       conf.level = 0.95) # alpha = 0.05

```

> c. If you were interested in the change in workers’ purchasing power from 1996 to 2015, would you use the results from (a) or (b)? Explain.

**Solution**

The results from part (b) adjust for changes in purchasing power. These results should be used.

> d. Using the data for 2015:
>      i. Construct a 95\% confidence interval for the mean of $AHE$ for high school graduates. 
>      ii. Construct a 95\% confidence interval for the mean of $AHE$ for
workers with a college degree.
>      iii. Construct a 95\% confidence interval for the difference between the
two means.

**Solution**

```{r}
# data: high school in 2015
CPS15hs <- CPS15[CPS15$bachelor == 0, ]

# data: college in 2015
CPS15col <- CPS15[CPS15$bachelor == 1, ]

# i. 95% CI for AHE in high school # ii. 95% CI for AHE in college
tapply(CPS15$ahe, CPS15$bachelor, E31a)

# iii. 95% CI for difference
t.test(CPS15hs$ahe, CPS15col$ahe, 
       alternative = c("two.sided"),
       mu = 0, # H0
       var.equal = FALSE,
       conf.level = 0.95) # alpha = 0.05

```


> e. Repeat (d) using the 1996 data expressed in \$2015.

**Solution**

```{r}
# data: high school in 2015
CPS96hs <- CPS96[CPS96$bachelor == 0, ]

# data: college in 2015
CPS96col <- CPS96[CPS96$bachelor == 1, ]

# i. 95% CI for AHE in high school # ii. 95% CI for AHE in college
tapply(CPS96$ahe_adjust, CPS96$bachelor, E31a)

# iii. 95% CI for difference
t.test(CPS96col$ahe_adjust, CPS96hs$ahe_adjust, 
       alternative = c("two.sided"),
       mu = 0, # H0
       var.equal = FALSE,
       conf.level = 0.95) # alpha = 0.05

```


> f. Using appropriate estimates, confidence intervals, and test statistics, answer the following questions:
>      i. Did real (inflation-adjusted) wages of high school graduates increase from 1996 to 2015?
>      ii. Did real wages of college graduates increase?
>      iii. Did the gap between earnings of college and high school graduates increase? Explain.

**Solution**

i. 
*  **Prepare**

$H_0: wage_{2015}^{hs} - wage_{1996}^{hs} = 0$ v.s. $H_1: wage_{2015}^{hs} - wage_{1996}^{hs} \neq 0$

Let the significance level be 0.05.


*  **Calculate**

Test statistics:
\begin{align}
    t 
    & = \frac{\left(\overline{w}_{2015}^{hs} - \overline{w}_{1996}^{hs}\right) - 0}{\sqrt{\frac{s_{2015,hs}^{2}}{n_{2015}^{hs}} + \frac{s_{1996,hs}^{2}}{n_{1996}^{hs}}}}
\end{align}

95\% confidence interval: 
\begin{align}
     \left(\overline{w}_{2015}^{hs} - \overline{w}_{1996}^{hs}\right) \pm Z_{0.025} \cdot \sqrt{\frac{s_{2015,hs}^{2}}{n_{2015}^{hs}} + \frac{s_{1996,hs}^{2}}{n_{1996}^{hs}}}
\end{align}

```{r}
E31fi <- function(x1, x2){
  # mean
  mu1 <- mean(x1); mu2 <- mean(x2)
  
  # s
  SD1 <- sd(x1); SD2 <- sd(x2)
  
  # n
  n1 <- length(x1); n2 <- length(x2)
  
  # difference in mean
  mu <- mu2 - mu1
  
  # difference in standard error
  se <- sqrt(SD1^2/n1 + SD2^2/n2)
  
  # 95% confidence interval
  lower <- round(mu - qnorm(0.975, mean = 0, sd = 1)*se, digit = 4)
  upper <- round(mu + qnorm(0.975, mean = 0, sd = 1)*se, digit = 4)
  CI <- paste(lower, "-"  ,upper)
  
  Table <- data.frame(mu, se, CI)
  colnames(Table) <- c("Mean", "Standard Error", "95% Confidence Interval")
  
  Table  
  
}

E31fi(CPS96hs$ahe_adjust, CPS15hs$ahe)

t.test(CPS15hs$ahe, CPS96hs$ahe_adjust, 
       alternative = c("two.sided"),
       mu = 0, # H0
       var.equal = FALSE,
       conf.level = 0.95) # alpha = 0.05

```


*  **Conclude**

Because $p-value > 0.05$, we do not reject $H_0$. There is no evidence that real (inflation-adjusted) wages of high school graduates increase from 1996 to 2015.

ii.

*  **Prepare**

$H_0: wage_{2015}^{col} - wage_{1996}^{col} = 0$ v.s. $H_1: wage_{2015}^{col} - wage_{1996}^{col} \neq 0$

Let the significance level be 0.05.


*  **Calculate**

Test statistics:
\begin{align}
    t 
    & = \frac{\left(\overline{w}_{2015}^{col} - \overline{w}_{1996}^{col}\right) - 0}{\sqrt{\frac{s_{2015,col}^{2}}{n_{2015}^{col}} + \frac{s_{1996,col}^{2}}{n_{1996}^{col}}}}
\end{align}

95\% confidence interval: 
\begin{align}
     \left(\overline{w}_{2015}^{col} - \overline{w}_{1996}^{col}\right) \pm Z_{0.025} \cdot \sqrt{\frac{s_{2015,col}^{2}}{n_{2015}^{col}} + \frac{s_{1996,col}^{2}}{n_{1996}^{col}}}
\end{align}

```{r}
E31fi(CPS96col$ahe_adjust, CPS15col$ahe)

t.test(CPS15col$ahe, CPS96col$ahe_adjust, 
       alternative = c("two.sided"),
       mu = 0, # H0
       var.equal = FALSE,
       conf.level = 0.95) # alpha = 0.05

```

*  **Conclude**

Because $p-value < 0.05$, we reject $H_0$. There is statistically significance evidence that real (inflation-adjusted) wages of college graduates increase from 1996 to 2015.


iii.

*  **Prepare**

$H_0: wage_{2015}^{diff} - wage_{1996}^{diff} = 0$ v.s. $H_1: wage_{2015}^{diff} - wage_{1996}^{diff} \neq 0$

Let the significance level be 0.05.


*  **Calculate**

Test statistics:
\begin{align}
    t 
    & = \frac{\left[ \left(\overline{w}_{2015}^{col} - \overline{w}_{2015}^{hs}\right) - \left(\overline{w}_{1996}^{col} - \overline{w}_{1996}^{hs}\right) \right] - 0}{\sqrt{ \left( \frac{s_{2015,col}^{2}}{n_{2015}^{col}} + \frac{s_{2015,hs}^{2}}{n_{2015}^{hs}} \right) + \left( \frac{s_{1996,col}^{2}}{n_{1996}^{col}} + \frac{s_{1996,hs}^{2}}{n_{1996}^{hs}} \right) }}
\end{align}

95\% confidence interval: 
\begin{align}
     & \left[ \left(\overline{w}_{2015}^{col} - \overline{w}_{2015}^{hs}\right) - \left(\overline{w}_{1996}^{col} - \overline{w}_{1996}^{hs}\right) \right] \pm Z_{0.025} \cdot \sqrt{ \left( \frac{s_{2015,col}^{2}}{n_{2015}^{col}} + \frac{s_{2015,hs}^{2}}{n_{2015}^{hs}} \right) + \left( \frac{s_{1996,col}^{2}}{n_{1996}^{col}} + \frac{s_{1996,hs}^{2}}{n_{1996}^{hs}} \right) }
\end{align}

```{r}
E31fiii <- function(x11, x12, x21, x22){
  # mean
  mu11 <- mean(x11); mu12 <- mean(x12)
  mu21 <- mean(x21); mu22 <- mean(x22)
  
  # mu
  mu1 <- mu11 - mu12
  mu2 <- mu21 - mu22
  
  # SE
  SD11 <- sd(x11); SD12 <- sd(x12)
  SD21 <- sd(x21); SD22 <- sd(x22)
  
  # n
  n11 <- length(x11); n12 <- length(x12)
  n21 <- length(x21); n22 <- length(x22)
  
  SD1 <- sqrt(SD11^2/n11 + SD12^2/n12)
  SD2 <- sqrt(SD21^2/n21 + SD22^2/n22)
  
  # difference in mean
  mu <- mu2 - mu1
  
  # difference in standard error
  se <- sqrt(SD1^2 + SD2^2)
  
  Table <- data.frame(mu, se, SD1, SD2)
  colnames(Table) <- c("Mean", "Standard Error", "SD1", "SD2")
  
  Table
}

E31fiii(CPS96col$ahe_adjust, CPS96hs$ahe_adjust, CPS15col$ahe, CPS15hs$ahe)
```


*  **Conclude**




> g. Table 3.1 presents information on the gender gap for college graduates. Prepare a similar table for high school graduates, using the 1996 and 2015 data. Are there any notable differences between the results for high school and college graduates?

**Solution**

```{r}
E31g <- function(x){
  # sample mean
  mu <- mean(x)
  
  # standard deviation
  SD <- sd(x)
  
  # numbers of sample
  n <- length(x)
  
  Table <- data.frame(mu, SD, n)
  
  Table
}

# data: male in 1996
CPS96hsm <- CPS96hs[CPS96hs$female == 0, ]

# data: female in 1996
CPS96hsf <- CPS96hs[CPS96hs$female == 1, ]

# statistics: male and female in 1996
tapply(CPS96hs$ahe_adjust, CPS96hs$female, E31g)

# difference in 1996
t.test(CPS96hsm$ahe_adjust, CPS96hsf$ahe_adjust, 
       alternative = c("two.sided"),
       mu = 0, # H0
       var.equal = FALSE,
       conf.level = 0.95) # alpha = 0.05

# statistics: male and female in 2015
tapply(CPS15hs$ahe, CPS15hs$female, E31g)

# data: male in 2015
CPS15hsm <- CPS15hs[CPS15hs$female == 0, ]

# data: female in 2015
CPS15hsf <- CPS15hs[CPS15hs$female == 1, ]

# difference in 2015
t.test(CPS15hsm$ahe, CPS15hsf$ahe, 
       alternative = c("two.sided"),
       mu = 0, # H0
       var.equal = FALSE,
       conf.level = 0.95) # alpha = 0.05

```



