---
title: "Empirical Exercise - E3.1"
author: "Chi-Yuan Fang"
date: "`r Sys.Date()`"
output:
  html_document:
    number_sections: yes
    toc: yes
  pdf_document:
    toc: yes

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#library(kableExtra)
```

```{r packages, include=FALSE}
library(openintro)
library(tidyverse)
data(COL)
```

> On the text website, <http://www.pearsonglobaleditions.com>, you will find the data file **CPS96_15**, which contains an extended version of the data set used in Table 3.1 of the text for the years 1996 and 2015. It contains data on full-time workers, ages 25–34, with a high school diploma or a B.A./B.S. as their highest degree. A detailed description is given in **CPS96_15_Description**, available on the website. Use these data to complete the following.

> a.
>      i. Compute the sample mean for average hourly earnings ($AHE$) in 1996 and 2015.
>      ii. Compute the sample standard deviation for $AHE$ in 1996 and 2015.
>      iii. Construct a 95\% confidence interval for the population means of $AHE$ in 1996 and 2015.
>      iv. Construct a 95\% confidence interval for the change in the population means of $AHE$ between 1996 and 2015.

**Solution**

```{r}
# import data 
library(readxl)
CPS96_15 <- read_xlsx("CPS96_15/CPS96_15.xlsx")

# data: 1996
CPS96 <- CPS96_15[CPS96_15$year == 1996,]

# data: 2015
CPS15 <- CPS96_15[CPS96_15$year == 2015,]

# i. sample mean # ii. standard error # iii. 95% CI
E31a <- function(x){
  # i. sample mean
  mu <- mean(x)
  
  # ii. standard error = sample standard deviation (s) / sqrt(n)
  se <- sd(x)/sqrt(length(x))
  
  # test
  test <- t.test(x, 
                 alternative = c("two.sided"),
                 mu = 0, # H0
                 conf.level = 0.95) # alpha = 0.05
  
  # iii. 95% confidence interval
  lower <- round(test$conf.int[1], digit = 4)
  upper <- round(test$conf.int[2], digit = 4)
  CI <- paste(lower, "-"  ,upper)
  
  Table <- data.frame(mu, se, CI)
  colnames(Table) <- c("Mean", "Standard Error", "95% Confidence Interval")
  
  Table

}

tapply(CPS96_15$ahe, CPS96_15$year, E31a)

# iv. 95% confidence interval for difference

t.test(CPS15$ahe, CPS96$ahe, 
       alternative = c("two.sided"),
       mu = 0, # H0
       var.equal = FALSE,
       conf.level = 0.95) # alpha = 0.05

```

> b. In 2015, the value of the Consumer Price Index (CPI) was 237.0. In 1996, the value of the CPI was 156.9. Repeat (a), but use AHE measured in real 2015 dollars (\$2015); that is, adjust the 1996 data for the price inflation that occurred between 1996 and 2015.

**Solution**

```{r}
# data: 1996
CPS96 <- CPS96_15[CPS96_15$year == 1996,]

# CPI
CPI_96 <- 156.9
CPI_15 <- 237

# adjusted 1996 AHE in $2015 
ahe_adjust <- CPS96$ahe * (CPI_15/CPI_96)

# data: 1996 including adjusted 1996 AHE in $2015 
CPS96 <- cbind(CPS96, ahe_adjust)

# i. sample mean
# ii. sample standard deviation
# iii. 95% confidence interval
# for adjusted 1996 AHE in $2015 
E31a(CPS96$ahe_adjust)

# iv. 95% confidence interval for difference
t.test(CPS15$ahe, CPS96$ahe_adjust, 
       alternative = c("two.sided"),
       mu = 0, # H0
       var.equal = FALSE,
       conf.level = 0.95) # alpha = 0.05

```

> c. If you were interested in the change in workers’ purchasing power from 1996 to 2015, would you use the results from (a) or (b)? Explain.

**Solution**

The results from part (b) adjust for changes in purchasing power. These results should be used.

> d. Using the data for 2015:
>      i. Construct a 95\% confidence interval for the mean of $AHE$ for high school graduates. 
>      ii. Construct a 95\% confidence interval for the mean of $AHE$ for
workers with a college degree.
>      iii. Construct a 95\% confidence interval for the difference between the
two means.

**Solution**

```{r}
# data: high school in 2015
CPS15hs <- CPS15[CPS15$bachelor == 0, ]

# data: college in 2015
CPS15col <- CPS15[CPS15$bachelor == 1, ]

# i. 95% CI for AHE in high school # ii. 95% CI for AHE in college
tapply(CPS15$ahe, CPS15$bachelor, E31a)

# iii. 95% CI for difference
t.test(CPS15hs$ahe, CPS15col$ahe, 
       alternative = c("two.sided"),
       mu = 0, # H0
       var.equal = FALSE,
       conf.level = 0.95) # alpha = 0.05

```


> e. Repeat (d) using the 1996 data expressed in \$2015.

**Solution**

```{r}
# data: high school in 2015
CPS96hs <- CPS96[CPS96$bachelor == 0, ]

# data: college in 2015
CPS96col <- CPS96[CPS96$bachelor == 1, ]

# i. 95% CI for AHE in high school # ii. 95% CI for AHE in college
tapply(CPS96$ahe_adjust, CPS96$bachelor, E31a)

# iii. 95% CI for difference
t.test(CPS96col$ahe_adjust, CPS96hs$ahe_adjust, 
       alternative = c("two.sided"),
       mu = 0, # H0
       var.equal = FALSE,
       conf.level = 0.95) # alpha = 0.05

```


> f. Using appropriate estimates, confidence intervals, and test statistics, answer the following questions:
>      i. Did real (inflation-adjusted) wages of high school graduates increase from 1996 to 2015?
>      ii. Did real wages of college graduates increase?
>      iii. Did the gap between earnings of college and high school graduates increase? Explain.

**Solution**

i. 
*  **Prepare**

$H_0: wage_{2015}^{hs} - wage_{1996}^{hs} = 0$ v.s. $H_1: wage_{2015}^{hs} - wage_{1996}^{hs} \neq 0$

Let the significance level be 0.05.


*  **Calculate**

Test statistics:
\begin{align}
    t 
    & = \frac{\left(\overline{w}_{2015}^{hs} - \overline{w}_{1996}^{hs}\right) - 0}{\sqrt{\frac{s_{2015,hs}^{2}}{n_{2015}^{hs}} + \frac{s_{1996,hs}^{2}}{n_{1996}^{hs}}}}
\end{align}

95\% confidence interval: 
\begin{align}
     \left(\overline{w}_{2015}^{hs} - \overline{w}_{1996}^{hs}\right) \pm Z_{0.025} \cdot \sqrt{\frac{s_{2015,hs}^{2}}{n_{2015}^{hs}} + \frac{s_{1996,hs}^{2}}{n_{1996}^{hs}}}
\end{align}

```{r}
E31fi <- function(x1, x2){
  # mean
  mu1 <- mean(x1); mu2 <- mean(x2)
  
  # s
  SD1 <- sd(x1); SD2 <- sd(x2)
  
  # n
  n1 <- length(x1); n2 <- length(x2)
  
  # difference in mean
  mu <- mu2 - mu1
  
  # difference in standard error
  se <- sqrt(SD1^2/n1 + SD2^2/n2)
  
  # 95% confidence interval
  lower <- round(mu - qnorm(0.975, mean = 0, sd = 1)*se, digit = 4)
  upper <- round(mu + qnorm(0.975, mean = 0, sd = 1)*se, digit = 4)
  CI <- paste(lower, "-"  ,upper)
  
  Table <- data.frame(mu, se, CI)
  colnames(Table) <- c("Mean", "Standard Error", "95% Confidence Interval")
  
  Table  
  
}

E31fi(CPS96hs$ahe_adjust, CPS15hs$ahe)

t.test(CPS15hs$ahe, CPS96hs$ahe_adjust, 
       alternative = c("two.sided"),
       mu = 0, # H0
       var.equal = FALSE,
       conf.level = 0.95) # alpha = 0.05

```


*  **Conclude**

Because $p-value > 0.05$, we do not reject $H_0$. There is no evidence that real (inflation-adjusted) wages of high school graduates increase from 1996 to 2015.

ii.

*  **Prepare**

$H_0: wage_{2015}^{col} - wage_{1996}^{col} = 0$ v.s. $H_1: wage_{2015}^{col} - wage_{1996}^{col} \neq 0$

Let the significance level be 0.05.


*  **Calculate**

Test statistics:
\begin{align}
    t 
    & = \frac{\left(\overline{w}_{2015}^{col} - \overline{w}_{1996}^{col}\right) - 0}{\sqrt{\frac{s_{2015,col}^{2}}{n_{2015}^{col}} + \frac{s_{1996,col}^{2}}{n_{1996}^{col}}}}
\end{align}

95\% confidence interval: 
\begin{align}
     \left(\overline{w}_{2015}^{col} - \overline{w}_{1996}^{col}\right) \pm Z_{0.025} \cdot \sqrt{\frac{s_{2015,col}^{2}}{n_{2015}^{col}} + \frac{s_{1996,col}^{2}}{n_{1996}^{col}}}
\end{align}

```{r}
E31fi(CPS96col$ahe_adjust, CPS15col$ahe)

t.test(CPS15col$ahe, CPS96col$ahe_adjust, 
       alternative = c("two.sided"),
       mu = 0, # H0
       var.equal = FALSE,
       conf.level = 0.95) # alpha = 0.05

```

*  **Conclude**

Because $p-value < 0.05$, we reject $H_0$. There is statistically significance evidence that real (inflation-adjusted) wages of college graduates increase from 1996 to 2015.


iii.

*  **Prepare**

$H_0: wage_{2015}^{diff} - wage_{1996}^{diff} = 0$ v.s. $H_1: wage_{2015}^{diff} - wage_{1996}^{diff} \neq 0$

Let the significance level be 0.05.


*  **Calculate**

Test statistics:
\begin{align}
    t 
    & = \frac{\left[ \left(\overline{w}_{2015}^{col} - \overline{w}_{2015}^{hs}\right) - \left(\overline{w}_{1996}^{col} - \overline{w}_{1996}^{hs}\right) \right] - 0}{\sqrt{ \left( \frac{s_{2015,col}^{2}}{n_{2015}^{col}} + \frac{s_{2015,hs}^{2}}{n_{2015}^{hs}} \right) + \left( \frac{s_{1996,col}^{2}}{n_{1996}^{col}} + \frac{s_{1996,hs}^{2}}{n_{1996}^{hs}} \right) }}
\end{align}

95\% confidence interval: 
\begin{align}
     & \left[ \left(\overline{w}_{2015}^{col} - \overline{w}_{2015}^{hs}\right) - \left(\overline{w}_{1996}^{col} - \overline{w}_{1996}^{hs}\right) \right] \pm Z_{0.025} \cdot \sqrt{ \left( \frac{s_{2015,col}^{2}}{n_{2015}^{col}} + \frac{s_{2015,hs}^{2}}{n_{2015}^{hs}} \right) + \left( \frac{s_{1996,col}^{2}}{n_{1996}^{col}} + \frac{s_{1996,hs}^{2}}{n_{1996}^{hs}} \right) }
\end{align}

```{r}
E31fiii <- function(x11, x12, x21, x22){
  # mean
  mu11 <- mean(x11); mu12 <- mean(x12)
  mu21 <- mean(x21); mu22 <- mean(x22)
  
  # mu
  mu1 <- mu11 - mu12
  mu2 <- mu21 - mu22
  
  # SE
  SD11 <- sd(x11); SD12 <- sd(x12)
  SD21 <- sd(x21); SD22 <- sd(x22)
  
  # n
  n11 <- length(x11); n12 <- length(x12)
  n21 <- length(x21); n22 <- length(x22)
  
  SD1 <- sqrt(SD11^2/n11 + SD12^2/n12)
  SD2 <- sqrt(SD21^2/n21 + SD22^2/n22)
  
  # difference in mean
  mu <- mu2 - mu1
  
  # difference in standard error
  se <- sqrt(SD1^2 + SD2^2)
  
  # 95% confidence interval
  lower <- round(mu - qnorm(0.975, mean = 0, sd = 1)*se, digit = 4)
  upper <- round(mu + qnorm(0.975, mean = 0, sd = 1)*se, digit = 4)
  CI <- paste(lower, "-"  ,upper)
  
  Table <- data.frame(mu, se, CI)
  colnames(Table) <- c("Mean", "Standard Error", "95% Confidence Interval")
  
  Table 
}

E31fiii(CPS96col$ahe_adjust, CPS96hs$ahe_adjust, CPS15col$ahe, CPS15hs$ahe)
```


*  **Conclude**

Because $0 \notin 95\%$ confidence interval, we reject $H_0$. There is statistically significance evidence that the gap between earnings of college and high school graduates increase.



> g. Table 3.1 presents information on the gender gap for college graduates. Prepare a similar table for high school graduates, using the 1996 and 2015 data. Are there any notable differences between the results for high school and college graduates?

**Solution**

```{r}
E31g <- function(x){
  # sample mean
  mu <- mean(x)
  
  # standard deviation
  SD <- sd(x)
  
  # numbers of sample
  n <- length(x)
  
  Table <- data.frame(mu, SD, n)
  
  Table
}

# data: male in 1996
CPS96hsm <- CPS96hs[CPS96hs$female == 0, ]

# data: female in 1996
CPS96hsf <- CPS96hs[CPS96hs$female == 1, ]

# statistics: male and female in 1996
tapply(CPS96hs$ahe_adjust, CPS96hs$female, E31g)

# difference in 1996
t.test(CPS96hsm$ahe_adjust, CPS96hsf$ahe_adjust, 
       alternative = c("two.sided"),
       mu = 0, # H0
       var.equal = FALSE,
       conf.level = 0.95) # alpha = 0.05

# statistics: male and female in 2015
tapply(CPS15hs$ahe, CPS15hs$female, E31g)

# data: male in 2015
CPS15hsm <- CPS15hs[CPS15hs$female == 0, ]

# data: female in 2015
CPS15hsf <- CPS15hs[CPS15hs$female == 1, ]

# difference in 2015
t.test(CPS15hsm$ahe, CPS15hsf$ahe, 
       alternative = c("two.sided"),
       mu = 0, # H0
       var.equal = FALSE,
       conf.level = 0.95) # alpha = 0.05

```


